### 2020-10-17 Paper/journal-article/journal-article.tex

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{An awesome journal} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Short Paper},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc header

\begin{document}
\begin{frontmatter}

  \title{Short Paper}
    \author[Some Institute of Technology]{Alice Anonymous\corref{1}}
   \ead{alice@example.com} 
    \author[Another University]{Bob Security}
   \ead{bob@example.com} 
    \author[Another University]{Cat Memes\corref{2}}
   \ead{cat@example.com} 
    \author[Some Institute of Technology]{Derek Zoolander\corref{2}}
   \ead{derek@example.com} 
      \address[Some Institute of Technology]{Department, Street, City, State, Zip}
    \address[Another University]{Department, Street, City, State, Zip}
      \cortext[1]{Corresponding Author}
    \cortext[2]{Equal contribution}

  \begin{abstract}
  This is the abstract.

  It consists of two paragraphs.
  \end{abstract}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(dplyr)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Attaching package: 'dplyr'
\end{verbatim}

\begin{verbatim}
## The following objects are masked from 'package:stats':
## 
##     filter, lag
\end{verbatim}

\begin{verbatim}
## The following objects are masked from 'package:base':
## 
##     intersect, setdiff, setequal, union
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(readr)}
\KeywordTok{library}\NormalTok{(tidyr)}
\KeywordTok{library}\NormalTok{(ggplot2)}
\KeywordTok{library}\NormalTok{(ggrepel)}

\CommentTok{# this is a lookup table matching MSOAs to major towns and cities}
\NormalTok{city_names <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{'../data-raw/Middle_Layer_Super_Output_Area__2011__to_Major_Towns_and_Cities__December_2015__Lookup_in_England_and_Wales.csv'}\NormalTok{) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Parsed with column specification:
## cols(
##   MSOA11CD = col_character(),
##   MSOA11NM = col_character(),
##   TCITY15CD = col_character(),
##   TCITY15NM = col_character(),
##   FID = col_double()
## )
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# change column name}
\NormalTok{city_names <-}\StringTok{ }\NormalTok{city_names }\OperatorTok{%>%}\StringTok{ }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{city =}\NormalTok{ TCITY15NM)}

\CommentTok{#unique cities}
\KeywordTok{unique}\NormalTok{(city_names}\OperatorTok{$}\NormalTok{city)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   [1] "London"               NA                     "Bolton"              
##   [4] "Stockport"            "Rochdale"             "Salford"             
##   [7] "Manchester"           "Oldham"               "Bury"                
##  [10] "Wigan"                "Rotherham"            "Liverpool"           
##  [13] "Doncaster"            "Sheffield"            "St Helens"           
##  [16] "Barnsley"             "Southport"            "Birkenhead"          
##  [19] "Birmingham"           "Coventry"             "Dudley"              
##  [22] "Newcastle upon Tyne"  "Sunderland"           "Sutton Coldfield"    
##  [25] "South Shields"        "West Bromwich"        "Gateshead"           
##  [28] "Leeds"                "Wolverhampton"        "Bradford"            
##  [31] "Walsall"              "Huddersfield"         "Wakefield"           
##  [34] "Solihull"             "Halifax"              "Leicester"           
##  [37] "Nottingham"           "Hartlepool"           "Derby"               
##  [40] "Blackpool"            "Grimsby"              "Warrington"          
##  [43] "Scunthorpe"           "Kingston upon Hull"   "York"                
##  [46] "Blackburn"            "Middlesbrough"        "Stockton-on-Tees"    
##  [49] "Darlington"           "Telford"              "Stoke-on-Trent"      
##  [52] "Swindon"              "Bristol"              "Peterborough"        
##  [55] "Luton"                "Southend-on-Sea"      "Gillingham"          
##  [58] "Plymouth"             "Weston-Super-Mare"    "Bath"                
##  [61] "Bournemouth"          "Poole"                "Portsmouth"          
##  [64] "Milton Keynes"        "Southampton"          "Bedford"             
##  [67] "Reading"              "Chatham"              "Bracknell"           
##  [70] "Slough"               "High Wycombe"         "Brighton and Hove"   
##  [73] "Cambridge"            "Carlisle"             "Chester"             
##  [76] "Chesterfield"         "Chelmsford"           "Basildon"            
##  [79] "Eastbourne"           "Hastings"             "Exeter"              
##  [82] "Colchester"           "Harlow"               "Cheltenham"          
##  [85] "Gloucester"           "Hemel Hempstead"      "St Albans"           
##  [88] "Stevenage"            "Basingstoke"          "Burnley"             
##  [91] "Watford"              "Maidstone"            "Preston"             
##  [94] "Norwich"              "Lincoln"              "Harrogate"           
##  [97] "Northampton"          "Newcastle-under-Lyme" "Oxford"              
## [100] "Mansfield"            "Shrewsbury"           "Burton upon Trent"   
## [103] "Woking"               "Guildford"            "Ipswich"             
## [106] "Crawley"              "Nuneaton"             "Worthing"            
## [109] "Redditch"             "Worcester"            "Swansea"             
## [112] "Cardiff"              "Newport"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# number of MSOAs in each city }
\NormalTok{no_msoas <-}\StringTok{ }\NormalTok{city_names }\OperatorTok{%>%}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{group_by}\NormalTok{(city) }\OperatorTok{%>%}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{tally}\NormalTok{()}

\CommentTok{##### CHOOSE YOU CITY }
\NormalTok{chosen_city <-}\StringTok{ "Manchester"}
\CommentTok{#create a directory to store data related to this city (does nothing if directory already exists)}
\KeywordTok{dir.create}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(}\StringTok{"../data/"}\NormalTok{, chosen_city), }\DataTypeTok{showWarnings =} \OtherTok{FALSE}\NormalTok{)}
\CommentTok{# create sub-directory to save plots as well}
\KeywordTok{dir.create}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(}\StringTok{"../data/"}\NormalTok{, chosen_city,}\StringTok{"/Plots"}\NormalTok{), }\DataTypeTok{showWarnings =} \OtherTok{FALSE}\NormalTok{)}

\CommentTok{##### CHOOSE YOU CITY }

\CommentTok{# flow data from the 2011 census https://www.nomisweb.co.uk/census/2011/bulk/rOD1}
\NormalTok{flows <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{'../data-raw/flow_data.csv'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Parsed with column specification:
## cols(
##   `Area of residence` = col_character(),
##   `Area of workplace` = col_character(),
##   `All categories: Method of travel to work` = col_double(),
##   `Work mainly at or from home` = col_double(),
##   `Underground, metro, light rail, tram` = col_double(),
##   Train = col_double(),
##   `Bus, minibus or coach` = col_double(),
##   Taxi = col_double(),
##   `Motorcycle, scooter or moped` = col_double(),
##   `Driving a car or van` = col_double(),
##   `Passenger in a car or van` = col_double(),
##   Bicycle = col_double(),
##   `On foot` = col_double(),
##   `Other method of travel to work` = col_double()
## )
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{###############}
\CommentTok{# MERGING NAMES WITH FLOW DATA (TO GET INTERNAL FLOWS IN ANY CITY)}
\CommentTok{###############}

\CommentTok{# add a column with the city name corresponding to each Residence MSOA}
\NormalTok{flows <-}\StringTok{ }\NormalTok{flows }\OperatorTok{%>%}\StringTok{ }\KeywordTok{left_join}\NormalTok{(city_names[,}\KeywordTok{c}\NormalTok{(}\StringTok{"MSOA11CD"}\NormalTok{, }\StringTok{"city"}\NormalTok{)],}
                             \DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"Area of residence"}\NormalTok{ =}\StringTok{ "MSOA11CD"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{city_origin =}\NormalTok{ city) }\CommentTok{# rename column so that we know it is referring to the 'Area of residence'}

\CommentTok{# add a column with the city name corresponding to each Workplace MSOA}
\NormalTok{flows <-}\StringTok{ }\NormalTok{flows }\OperatorTok{%>%}\StringTok{ }\KeywordTok{left_join}\NormalTok{(city_names[,}\KeywordTok{c}\NormalTok{(}\StringTok{"MSOA11CD"}\NormalTok{, }\StringTok{"city"}\NormalTok{)],}
                             \DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"Area of workplace"}\NormalTok{ =}\StringTok{ "MSOA11CD"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{city_dest =}\NormalTok{ city) }\CommentTok{# rename column so that we know it is referring to the 'Area of workplace'}

\CommentTok{# get mode share for all cities - just for report}
\NormalTok{cycle_mode_share <-}\StringTok{ }\NormalTok{flows }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(city_origin }\OperatorTok{==}\StringTok{ }\NormalTok{city_dest) }\OperatorTok{%>%}\StringTok{   }\CommentTok{# only internal flows}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(city_origin, city_dest) }\OperatorTok{%>%}\StringTok{   }
\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{mode_share =}\NormalTok{ (}\KeywordTok{sum}\NormalTok{(Bicycle) }\OperatorTok{/}\StringTok{ }\KeywordTok{sum}\NormalTok{(}\StringTok{`}\DataTypeTok{All categories: Method of travel to work}\StringTok{`}\NormalTok{)) }\OperatorTok{*}\DecValTok{100}\NormalTok{,}
            \DataTypeTok{all =} \KeywordTok{sum}\NormalTok{(}\StringTok{`}\DataTypeTok{All categories: Method of travel to work}\StringTok{`}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{arrange}\NormalTok{(mode_share)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `summarise()` regrouping output by 'city_origin' (override with `.groups` argument)
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# plot }
\NormalTok{cycle_mode_share }\OperatorTok{%>%}\StringTok{ }\CommentTok{#filter(city_origin != 'London') %>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(all, mode_share)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{alpha =}\NormalTok{ mode_share), }\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\CommentTok{# some filtering labels for aesthetic purposes. Add some cities explicitly as they are in my case study}
\StringTok{  }\KeywordTok{geom_label_repel}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =} \KeywordTok{ifelse}\NormalTok{(mode_share}\OperatorTok{>}\DecValTok{12} \OperatorTok{|}\StringTok{ }\NormalTok{all}\OperatorTok{>}\StringTok{ }\NormalTok{(}\FloatTok{2.3}\OperatorTok{*}\KeywordTok{mean}\NormalTok{(all)) }\OperatorTok{|}\StringTok{ }\NormalTok{city_origin }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{'Manchester'}\NormalTok{, }\StringTok{'Nottingham'}\NormalTok{, }\StringTok{'Birmingham'}\NormalTok{) , }\KeywordTok{as.character}\NormalTok{(city_origin),}\StringTok{''}\NormalTok{)), }
                       \DataTypeTok{size =}\FloatTok{2.5}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x=}\StringTok{"Total Number of Commuters"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"Cycling Mode Share (%)"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{trans=}\StringTok{'log10'}\NormalTok{, }\DataTypeTok{labels =}\NormalTok{ scales}\OperatorTok{::}\NormalTok{comma) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_minimal}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{journal-article_files/figure-latex/unnamed-chunk-1-1.pdf}

\hypertarget{what-affects-the-decision-to-cycle}{%
\section{What Affects the Decision To
Cycle}\label{what-affects-the-decision-to-cycle}}

\hypertarget{planning-cycling-networks}{%
\section{Planning Cycling Networks}\label{planning-cycling-networks}}

\hypertarget{potential-cycling-demand}{%
\section{Potential Cycling Demand}\label{potential-cycling-demand}}

\begin{align}\label{eq:pcteqn}
     logit(C_{p}) = & -4.018 - 0.6369d +  1.988\sqrt{d} + 0.008775d^2\\ & - 0.2555s + 0.00206ds -0.1234\sqrt{d}s\nonumber 
\end{align}

\hypertarget{routing}{%
\section{Routing}\label{routing}}

\hypertarget{road-segment-prioritization}{%
\section{Road Segment
Prioritization}\label{road-segment-prioritization}}

\hypertarget{analysis-of-methodology}{%
\section{Analysis of Methodology}\label{analysis-of-methodology}}

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\end{document}