### 2021-03-13 appendix.Rmd
Update renv

---
title: "Appendix"
author: "Hussein Mahfouz, Robin Lovelace, Elsa Arcaute"
date: "13/03/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reproducing the results for other cities

To reproduce the results presented for Manchester for other cities there are a few computing pre-requisites:

- A reasonably powerful computer, e.g. with 8 GB RAM
- The statistical programming language R installed (we recommend running it in the RStudio editor)
- The source code of the paper repository, which can be download from `url TBC`

```{r, eval=FALSE, echo=FALSE}
required_packages = renv::dependencies()
renv::snapshot()
```

# Birmingham

# Leeds

# Nottingham