### 2021-03-13 appendix.Rmd
Update appendix - working

---
title: "Appendix"
author: "Hussein Mahfouz, Robin Lovelace, Elsa Arcaute"
date: "13/03/2021"
output:
  pdf_document:
    extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reproducing the results for other cities

To reproduce the results presented for Manchester for other cities there are a few computing pre-requisites:

- A reasonably powerful computer, e.g. with 8 GB RAM
- The statistical programming language R installed (we recommend running it in the RStudio editor)
- The source code of the paper repository, which can be download from `url TBC`

```{r, eval=FALSE, echo=FALSE}
# run these lines to update the snapshot
required_packages = renv::dependencies()
renv::snapshot()
```

To ensure you have the necessary packages installed, open the `cycle-networks.Rproj` file in RStudio and run the following commands to ensure you have the necessary packages installed:

```{r, eval=FALSE}
install.packages("renv")
renv::restore()
```

The `build.R` script contains all of the necessary steps to reproduce the results for any city in the UK.
Open this file with the following command and run the code line-by-line.

```{r, eval=FALSE}
file.edit("R/build.R")
```

<!-- Todo: make the .zip download work for anyone, e.g. by uploading the data to the osf.io website alongside the preprint. -->

You can select a city to build by executing the following command:

```{r}
chosen_city <- "Leeds"
```

Replacing `Leeds` with any other city will make the script work for that city.
Note: this only works for Local Authority names in England and Wales listed below.

In this script, you choose which city you wish to run the analysis from
this list of available towns and cities:

    ##   [1] "Barnsley"             "Basildon"             "Basingstoke"         
    ##   [4] "Bath"                 "Bedford"              "Birkenhead"          
    ##   [7] "Birmingham"           "Blackburn"            "Blackpool"           
    ##  [10] "Bolton"               "Bournemouth"          "Bracknell"           
    ##  [13] "Bradford"             "Brighton and Hove"    "Bristol"             
    ##  [16] "Burnley"              "Burton upon Trent"    "Bury"                
    ##  [19] "Cambridge"            "Cardiff"              "Carlisle"            
    ##  [22] "Chatham"              "Chelmsford"           "Cheltenham"          
    ##  [25] "Chester"              "Chesterfield"         "Colchester"          
    ##  [28] "Coventry"             "Crawley"              "Darlington"          
    ##  [31] "Derby"                "Doncaster"            "Dudley"              
    ##  [34] "Eastbourne"           "Exeter"               "Gateshead"           
    ##  [37] "Gillingham"           "Gloucester"           "Grimsby"             
    ##  [40] "Guildford"            "Halifax"              "Harlow"              
    ##  [43] "Harrogate"            "Hartlepool"           "Hastings"            
    ##  [46] "Hemel Hempstead"      "High Wycombe"         "Huddersfield"        
    ##  [49] "Ipswich"              "Kingston upon Hull"   "Leeds"               
    ##  [52] "Leicester"            "Lincoln"              "Liverpool"           
    ##  [55] "London"               "Luton"                "Maidstone"           
    ##  [58] "Manchester"           "Mansfield"            "Middlesbrough"       
    ##  [61] "Milton Keynes"        "Newcastle upon Tyne"  "Newcastle-under-Lyme"
    ##  [64] "Newport"              "Northampton"          "Norwich"             
    ##  [67] "Nottingham"           "Nuneaton"             "Oldham"              
    ##  [70] "Oxford"               "Peterborough"         "Plymouth"            
    ##  [73] "Poole"                "Portsmouth"           "Preston"             
    ##  [76] "Reading"              "Redditch"             "Rochdale"            
    ##  [79] "Rotherham"            "Salford"              "Scunthorpe"          
    ##  [82] "Sheffield"            "Shrewsbury"           "Slough"              
    ##  [85] "Solihull"             "South Shields"        "Southampton"         
    ##  [88] "Southend-on-Sea"      "Southport"            "St Albans"           
    ##  [91] "St Helens"            "Stevenage"            "Stockport"           
    ##  [94] "Stockton-on-Tees"     "Stoke-on-Trent"       "Sunderland"          
    ##  [97] "Sutton Coldfield"     "Swansea"              "Swindon"             
    ## [100] "Telford"              "Wakefield"            "Walsall"             
    ## [103] "Warrington"           "Watford"              "West Bromwich"       
    ## [106] "Weston-Super-Mare"    "Wigan"                "Woking"              
    ## [109] "Wolverhampton"        "Worcester"            "Worthing"            
    ## [112] "York"

# Birmingham

\subsubsection{Potential Demand}

\begin{figure} [H] 
    \centering % puts the image in the horizontal centre of the page
    \captionsetup{font=footnotesize,labelfont=footnotesize}
    \includegraphics[width=.4\textwidth]{data/Birmingham/Plots/mode_share_increase_vs_performance_smooth_Birmingham} 
    \caption*{Distribution of Cycling Demand Relative to OD Pair Performance} 
\end{figure}

\begin{figure} [H] 
    \centering % puts the image in the horizontal centre of the page
    \captionsetup{font=footnotesize,labelfont=footnotesize}
    \includegraphics[width=.85\textwidth]{data/Birmingham/Plots/desire_facet_cycling_Birmingham} 
    \caption*{Current and Potential Cycling Flow (Birmingham)} 
\end{figure}

\newpage

\subsubsection{Routing Cycling Flows}

\begin{figure} [H] 
    \centering % puts the image in the horizontal centre of the page
    \captionsetup{font=footnotesize,labelfont=footnotesize}
    \includegraphics[width=.9\textwidth]{data/Birmingham/Plots/flows_facet_unweighted_Birmingham} 
    \caption*{Flow Results Based on \textbf{Unweighted} Shortest Paths (Birmingham)} 
\end{figure}

\begin{figure} [H] 
    \centering % puts the image in the horizontal centre of the page
    \captionsetup{font=footnotesize,labelfont=footnotesize}
    \includegraphics[width=.9\textwidth]{data/Birmingham/Plots/flows_facet_weighted_Birmingham} 
    \caption*{Flow Results Based on \textbf{Weighted} Shortest Paths (Birmingham)} 
\end{figure}

\subsubsection{Algorithm 2 (Utilitarian)}

\begin{figure} [H] 
\centering
\captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions
\begin{subfigure}{.4\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{data/Birmingham/Plots/Growth_Results/growth_existing_infra_satisfied_km_all_flow_column_Birmingham}
  \caption{\% of Person-Km Satisfied}
\end{subfigure}
\begin{subfigure}{.4\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{data/Birmingham/Plots/Growth_Results/growth_existing_infra_satisfied_km_community_flow_column_Birmingham}
  \caption{\% of Person-Km Satisfied in Each Community}
\end{subfigure}
\caption*{Effectiveness of Alg. 2 on Overall and Community-Wide Person-Km Satisfied (Birmingham)}
\end{figure}

\subsubsection{Algorithm 3 (Egalitarian)}

\begin{figure} [H] 
\centering
\captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions
\begin{subfigure}{.4\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{data/Birmingham/Plots/Growth_Results/growth_community_4_satisfied_km_all_flow_column_Birmingham}
  \caption{\% of Person-Km Satisfied}
\end{subfigure}
\begin{subfigure}{.4\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{data/Birmingham/Plots/Growth_Results/growth_community_4_satisfied_km_community_flow_column_Birmingham}
  \caption{\% of Person-Km Satisfied in Each Community}
\end{subfigure}
\caption*{Effectiveness of Alg. 3 on Overall and Community-Wide Person-Km Satisfied (Birmingham)}
\end{figure}

\begin{figure} [H] 
\centering
\captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions
\includegraphics[width=1\linewidth]{data/Birmingham/Plots/Growth_Results/growth_community_4_priority_all_FLOW_Birmingham}
\caption*{Results of Alg. 3 (Birmingham)}
\end{figure}

# Leeds

\subsubsection{Potential Demand}

\begin{figure} [H] 
    \centering % puts the image in the horizontal centre of the page
    \captionsetup{font=footnotesize,labelfont=footnotesize}
    \includegraphics[width=.4\textwidth]{data/Leeds/Plots/mode_share_increase_vs_performance_smooth_Leeds} 
    \caption*{Distribution of Cycling Demand Relative to OD Pair Performance} 
\end{figure}

<!-- \begin{figure} [H]  -->
<!--     \centering % puts the image in the horizontal centre of the page -->
<!--     \captionsetup{font=footnotesize,labelfont=footnotesize} -->
<!--     \includegraphics[width=.85\textwidth]{data/Leeds/Plots/desire_facet_cycling_Leeds}  -->
<!--     \caption*{Current and Potential Cycling Flow (Leeds)}  -->
<!-- \end{figure} -->

\newpage

\subsubsection{Routing Cycling Flows}

\begin{figure} [H] 
    \centering % puts the image in the horizontal centre of the page
    \captionsetup{font=footnotesize,labelfont=footnotesize}
    \includegraphics[width=.9\textwidth]{data/Leeds/Plots/flows_facet_unweighted_Leeds} 
    \caption*{Flow Results Based on \textbf{Unweighted} Shortest Paths (Leeds)} 
\end{figure}

\begin{figure} [H] 
    \centering % puts the image in the horizontal centre of the page
    \captionsetup{font=footnotesize,labelfont=footnotesize}
    \includegraphics[width=.9\textwidth]{data/Leeds/Plots/flows_facet_weighted_Leeds} 
    \caption*{Flow Results Based on \textbf{Weighted} Shortest Paths (Leeds)} 
\end{figure}

\subsubsection{Algorithm 2 (Utilitarian)}

<!-- \begin{figure} [H]  -->
<!-- \centering -->
<!-- \captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions -->
<!-- \begin{subfigure}{.4\textwidth} -->
<!--   \centering -->
<!--   \includegraphics[width=1\linewidth]{data/Leeds/Plots/Growth_Results/growth_existing_infra_satisfied_km_all_flow_column_Leeds} -->
<!--   \caption{\% of Person-Km Satisfied} -->
<!-- \end{subfigure} -->
<!-- \begin{subfigure}{.4\textwidth} -->
<!--   \centering -->
<!--   \includegraphics[width=1\linewidth]{data/Leeds/Plots/Growth_Results/growth_existing_infra_satisfied_km_community_flow_column_Leeds} -->
<!--   \caption{\% of Person-Km Satisfied in Each Community} -->
<!-- \end{subfigure} -->
<!-- \caption*{Effectiveness of Alg. 2 on Overall and Community-Wide Person-Km Satisfied (Leeds)} -->
<!-- \end{figure} -->

\subsubsection{Algorithm 3 (Egalitarian)}

\begin{figure} [H]
\centering
\captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions
\begin{subfigure}{.4\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{data/Leeds/Plots/Growth_Results/growth_utilitarian_satisfied_km_all_flow_column}
  \caption{\% of Person-Km Satisfied}
\end{subfigure}
\begin{subfigure}{.4\textwidth}
  \centering
  \includegraphics[width=1\linewidth]{data/Leeds/Plots/Growth_Results/growth_utilitarian_satisfied_km_community_flow_column}
  \caption{\% of Person-Km Satisfied in Each Community}
\end{subfigure}
\caption*{Effectiveness of Alg. 3 on Overall and Community-Wide Person-Km Satisfied (Leeds)}
\end{figure}

\begin{figure} [H] 
\centering
\captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions
\includegraphics[width=1\linewidth]{data/Leeds/Plots/Growth_Results/growth_egalitarian_satisfied_km_community_flow_column}
\caption*{Results of Alg. 3 (Leeds)}
\end{figure}

# Nottingham